---
title: "Hand Sanitizer"
author: "Tim Hogan"
date: "3/6/2020"
output: ioislides
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Context

- After hand sanitizer was installed, nurses underwent a training program to use hand sanitizers.
- Before the program, in 51 visits, hand sanitizer was used 13 times.
- After two months, when revisiting the nurses, in 56 visits, 50 of them involved hand sanitizer.
- Can we measure how they improved?

## How much did they improve?

- We can approximate this by calculating p, the probability a nurse will use hand sanitizer.
- In the first study, the proportion of nurses that used hand sanitizer was 13/51, about ~25.4%.
- After the study, the proportion was 50/56, which is about 89%.
- However, we can do better. Using a Bayesian approach, we can find a probability that incorporates both experiments

## Incorporating Bayes' Law

- In this case, we're looking at a problem with a fixed number of studies, a success case, and a failure case.
- We can use a binomial distribution, with parameters n, the number of cases, and p, the probability of success. We also now X, the number of successes, or out data.
- We can look closer using Bayes Law, which compares our data, X, our likelihood, X given p (or $P(X|P)$) and our prior information for our probability, $P(p)$.

$$ P(p|X) \propto P(X|p)*P(p) $$
- So, to move forward, we need to find our likelihood and prior information.

## Conjugate Priors

- The likelihood of a binomial distribution is $P(X|p) \propto p^{x}(1-p)^{n-x}$.
- We can incorporate prior information by using conjugate priors. This allows us to incorporate prior information in a very easy manner.
- First, we can take terms with data in our likelihood and turn them into parameters. Our data is x and n, since p is an underlying component. 
- This gives us a distribution $P(p) \propto p^{\alpha-1}(1-p)^{\beta-1}$. This is the beta distribution with parameters $\alpha$ and $\beta$, the conjugate prior for the binomial distribution!

## Prior to Posterior

- To find our posterior, we can refer back to Bayes Law, $P(p|X) \propto P(X|p)*P(p)$.
- We can plug in our likelihood and posterior to find that
$$ P(p|X) \propto p^{x}(1-p)^{n-x} p^{\alpha-1}(1-p)^{\beta-1}$$
$$ P(p|X) \propto p^{x+\alpha-1}(1-p)^{n-x+\beta-1}$$

- This, once again, is a beta distribution, $p^{a-1} + (1-p)^{b-1}$ with paramters $a = x+ \alpha$ and $b =n-x+\beta$. In other words, it's comparing the number of successes and failures, adjusted by prior information parameters $alpha$ and $beta$.

## An uninformative prior. 

- Suppose we have no prior information. Our prior can simply be a flat line, which is a beta distribution with parameters $\alpha=1$ and $\beta=1$.
- We also know for our experiment that x is the 50 people that used hand sanitizer and n is 56 study cases. Plugging in, we can find our posterior will follow a beta with parameters $a = x + \alpha = 50+1=51$ and $b =n-x+\beta= 56-50=7$.
- This distribution is a beta, so it has a mean of $\frac{\alpha}{\alpha+\beta} = \frac{51}{51+7}= 0.88$, and mode $\frac{\alpha-1}{\alpha+\beta-2} = \frac{51-1}{51+7-2} = 0.89$.
- Because we know this distribution, we can simulat it in R using a beta with parameters $a = 50$ and $b = 6$. This graph below shows the distribution of 1000 sampled cases.


## Simulating posterior with uninformative prior


```{r, echo=F}
set.seed(5)
x = 50
n = 56
a1 = 1
b1 = 1
sim1 <- rbeta(1000,x+a1,n-x+b1)
hist(sim1,main="1000 Simulated Cases with uninformative prior",xlab="Probability")

```

## Simulating posterior with uninformative prior

- This simulation has a mean `r mean(sim1)`, which is very close to our estimated mean 0.88!

```{r,echo=F}
hist(sim1,main="1000 Simulated Cases with uninformative prior",xlab="Probability")
```

## Incorporating past study into prior

- We can also use the initial study's data as prior information.
- In their study, they had 13 successes out of 51 cases. This can be approximated with prior parameters $\alpha=13$ and $\beta = 51-13=38$.
- This would lead to a conjugate beta with parameters $a = 50 + 13 = 63$ and $b = 6 + 38 = 44$. This distribution would have a mean of $\frac{63}{63+44} = 0.59$, and a mode of $\frac{63-1}{63+44-2} =0.59$. 
- Like last time, we can make 1000 simulations from a beta distribution in R with our $a$ and $b$ parametres.

## Simulating Posterior with Information
```{r,echo=F}
set.seed(6)
a2 = 13
b2 = 38
sim2 <- rbeta(1000,x+a2,n-x+b2)
```

- Our simulation has a mean `r mean(sim2)`, which is close to our estimated mean of 0.59.

```{r, echo=F}
hist(sim2,main="1000 Simulated Cases with informative prior",xlab="Probability")
```

## Conclusions.

- We analyzed the probability of nurses using hand sanitizer from these problems using an uniformative and informative case with Bayesian likelihood and conjugate priors.
- With no prior information, we estimated a mean probability of 0.88. With  information from the past experiment, we estimated a mean probability 0.59.
- Both of these distributions show a major improvement from the original rate of 0.25. However, the "informed" prior is lower, since it considers the past experiment as well as the most recent one.
- As successful cases and studies continue, the posterior can continue to be updated to incorporate both past data and current data.

